---
title: "Untitled"
author: "Nacima Abdi"
date: "05/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Loading Packages
```{r}
library(VennDiagram)
require(gridExtra)

```

Comparing Tolerant and susceptible DGE results from the early response
```{r}

#Load upregulated DGE results
T1_sus_up <- read.csv("Upregulated_genes_susceptible_T1_Root.csv")
##number of genes upregulated at T1 for susceptible
size_T1_sus_up <- nrow(T1_sus_up)
head(size_T1_sus_up)

T1_tol_up <-read.csv("Upregulated_genes_tolerant_T1_Root.csv")
##number of genes upregulated at T1 for tolerant
size_T1_tol_up <- nrow(T1_tol_up)
head(size_T1_tol_up)

#upregulated genes shared between tolerant and susceptible at T1
T1_up_genes_overlap = intersect(T1_sus_up$susR_T1_Up_genes, T1_tol_up$tolR_T1_Up_genes)
write.csv(data.frame(T1_up_genes_overlap), file= "Upregulated_genes_common_T1_Root.csv", row.names = FALSE)
##number of upregulated genes shared between tolerant and susceptible at T1
size_T1_up <- length(T1_up_genes_overlap)
head(size_T1_up)


#down
#Load downregulated DGE results
T1_sus_down <- read.csv("Downregulated_genes_susceptible_T1_Root.csv")
##number of genes downregulated at T1 for susceptible
size_T1_sus_down <- nrow(T1_sus_down)
head(size_T1_sus_down)

#Load downregulated DGE results
T1_tol_down <-read.csv("Downregulated_genes_tolerant_T1_Root.csv")
##number of genes downregulated at T1 for tolerant
size_T1_tol_down <- nrow(T1_tol_down)
head(size_T1_tol_down)

#downregulated genes shared between tolerant and susceptible at T1
T1_down_genes_overlap = intersect(T1_sus_down$susR_T1_Down_genes, T1_tol_down$tolR_T1_Down_genes)
write.csv(data.frame(T1_down_genes_overlap), file= "Downregulated_genes_common_T1_Root.csv", row.names = FALSE)
#number of upregulated genes shared between tolerant and susceptible at T1
size_T1_down <- length(T1_down_genes_overlap)
head(size_T1_down)

```
Venn diagriam for upregulated genes at T1
```{r}
#ven diagram

venn.plot <- draw.pairwise.venn(
	area1 = size_T1_tol_up,
	area2 = size_T1_sus_up,
	cross.area = size_T1_up,
	category = c("Drought Tolerant ", "Drought Susceptible"),
	fill = c("blue", "red"),
	euler.d = TRUE, 
	scaled = TRUE,
	lty = "blank",
	cex = 2,
	cat.dist = -0.08,
	cat.cex = 1.5,
	ext.line.lty = "dashed"
	);


venn.plot <- grid.arrange(gTree(children=venn.plot), top=textGrob("Upregulated Genes at Early Response", gp=gpar(fontsize=25,font=8)) )

# Writing to file
tiff(filename = "Upregulated Genes T1 Root Venn2_diagram.tiff");
grid.draw(venn.plot);
dev.off()
```


Venn diagriam for downregulated genes at T1
```{r}
#ven diagram

venn.plot <- draw.pairwise.venn(
	area1 = size_T1_tol_down,
	area2 = size_T1_sus_down,
	cross.area = size_T1_down,
	category = c("Drought Tolerant","Drought Susceptible"),
	fill = c("blue", "red"),
	euler.d = TRUE, 
	scaled = TRUE,
	lty = "blank",
	cex = 2,
	cat.dist = -0.08,
	cat.cex = 1.5,
	ext.line.lty = "dashed"
	);


venn.plot <- grid.arrange(gTree(children=venn.plot), top=textGrob("Downregulated Genes at Early Response", gp=gpar(fontsize=25,font=8)) )

# Writing to file
tiff(filename = "Downregulated Genes T1 Root Venn2_diagram.tiff");
grid.draw(venn.plot);
dev.off()
```


Comparing Tolerant and susceptible DGE results from the Late response
```{r}

#Load upregulated DGE results
T2_sus_up <- read.csv("Upregulated_genes_susceptible_T2_Root.csv")
##number of genes upregulated at T1 for susceptible
size_T2_sus_up <- nrow(T2_sus_up)
head(size_T2_sus_up)

T2_tol_up <-read.csv("Upregulated_genes_tolerant_T2_Root.csv")
##number of genes upregulated at T2 for tolerant
size_T2_tol_up <- nrow(T2_tol_up)
head(size_T2_tol_up)

#upregulated genes shared between tolerant and susceptible at T2
T2_up_genes_overlap = intersect(T2_sus_up$susR_T2_Up_genes, T2_tol_up$tolR_T2_Up_genes)
write.csv(data.frame(T2_up_genes_overlap), file= "Upregulated_genes_common_T2_Root.csv", row.names = FALSE)
##number of upregulated genes shared between tolerant and susceptible at T2
size_T2_up <- length(T2_up_genes_overlap)
head(size_T2_up)


#down
#Load downregulated DGE results
T2_sus_down <- read.csv("Downregulated_genes_susceptible_T2_Root.csv")
##number of genes downregulated at T2 for susceptible
size_T2_sus_down <- nrow(T2_sus_down)
head(size_T2_sus_down)

#Load downregulated DGE results
T2_tol_down <-read.csv("Downregulated_genes_tolerant_T2_Root.csv")
##number of genes downregulated at T2 for tolerant
size_T2_tol_down <- nrow(T2_tol_down)
head(size_T2_tol_down)

#downregulated genes shared between tolerant and susceptible at T2
T2_down_genes_overlap = intersect(T2_sus_down$susR_T2_Down_genes, T2_tol_down$tolR_T2_Down_genes)
write.csv(data.frame(T2_down_genes_overlap), file= "Downregulated_genes_common_T2_Root.csv", row.names = FALSE)
##number of upregulated genes shared between tolerant and susceptible at T2
size_T2_down <- length(T2_down_genes_overlap)
head(size_T2_down)

```
Venn diagriam for upregulated genes at T2
```{r}
#ven diagram

venn.plot <- draw.pairwise.venn(
	area1 = size_T2_sus_up,
	area2 = size_T2_tol_up,
	cross.area = size_T2_up,
	category = c("Drought Susceptible","Drought Tolerant "),
	fill = c("blue", "red"),
	euler.d = TRUE, 
	scaled = TRUE,
	lty = "blank",
	cex = 2,
	cat.dist = -0.03,
	cat.cex = 1.5,
	ext.line.lty = "dashed"
	);


venn.plot <- grid.arrange(gTree(children=venn.plot), top=textGrob("Upregulated Genes at Late Response", gp=gpar(fontsize=25,font=8)) )

# Writing to file
tiff(filename = "Upregulated Genes T2 Root Venn2_diagram.tiff");
grid.draw(venn.plot);
dev.off()
```


Venn diagriam for downregulated genes at T2
```{r}
#ven diagram

venn.plot <- draw.pairwise.venn(
	area1 = size_T2_sus_down,
	area2 = size_T2_tol_down,
	cross.area = size_T2_down,
	category = c("Drought Susceptible","Drought Tolerant "),
	fill = c("blue", "red"),
	euler.d = TRUE, 
	scaled = TRUE,
	lty = "blank",
	cex = 2,
	cat.dist = -0.05,
	cat.cex = 1.5,
	ext.line.lty = "dashed"
	);


venn.plot <- grid.arrange(gTree(children=venn.plot), top=textGrob("Downregulated Genes at Late Response", gp=gpar(fontsize=25,font=8)) )

# Writing to file
tiff(filename = "Downregulated Genes T2 Root Venn2_diagram.tiff");
grid.draw(venn.plot);
dev.off()
```

Comparing Tolerant and susceptible DGE results from Recovery
```{r}


T3_sus_up <- read.csv("Upregulated_genes_susceptible_T3_Root.csv")
##number of genes upregulated at T1 for susceptible
size_T3_sus_up <- nrow(T3_sus_up)
head(size_T3_sus_up)

T3_tol_up <-read.csv("Upregulated_genes_tolerant_T3_Root.csv")
##number of genes upregulated at T3 for tolerant
size_T3_tol_up <- nrow(T3_tol_up)
head(size_T3_tol_up)

#upregulated genes shared between tolerant and susceptible at T3
T3_up_genes_overlap = intersect(T3_sus_up$susR_T3_Up_genes, T3_tol_up$tolR_T3_Up_genes)
write.csv(data.frame(T3_up_genes_overlap), file= "Upregulated_genes_common_T3_Root.csv", row.names = FALSE)
##number of upregulated genes shared between tolerant and susceptible at T3
size_T3_up <- length(T3_up_genes_overlap)
head(size_T3_up)


#down
#Load downregulated DGE results
T3_sus_down <- read.csv("Downregulated_genes_susceptible_T3_Root.csv")
##number of genes downregulated at T3 for susceptible
size_T3_sus_down <- nrow(T3_sus_down)
head(size_T3_sus_down)

#Load downregulated DGE results
T3_tol_down <-read.csv("Downregulated_genes_tolerant_T3_Root.csv")
##number of genes downregulated at T3 for tolerant
size_T3_tol_down <- nrow(T3_tol_down)
head(size_T3_tol_down)

#downregulated genes shared between tolerant and susceptible at T3
T3_down_genes_overlap = intersect(T3_sus_down$susR_T3_Down_genes, T3_tol_down$tolR_T3_Down_genes)
write.csv(data.frame(T3_down_genes_overlap), file= "Downregulated_genes_common_T3_Root.csv", row.names = FALSE)
##number of upregulated genes shared between tolerant and susceptible at T3
size_T3_down <- length(T3_down_genes_overlap)
head(size_T3_down)

```
Venn diagriam for upregulated genes at T3
```{r}
#ven diagram

venn.plot <- draw.pairwise.venn(
	area1 = size_T3_sus_up,
	area2 = size_T3_tol_up,
	cross.area = size_T3_up,
	category = c("Drought Susceptible","Drought Tolerant "),
	fill = c("blue", "red"),
	euler.d = TRUE, 
	scaled = TRUE,
	lty = "blank",
	cex = 2,
	cat.dist = -0.02,
	cat.cex = 1.5,
	ext.line.lty = "dashed"
	);


venn.plot <- grid.arrange(gTree(children=venn.plot), top=textGrob("Upregulated Genes at Recovery", gp=gpar(fontsize=25,font=8)) )

# Writing to file
tiff(filename = "Upregulated Genes T3 Root Venn2_diagram.tiff");
grid.draw(venn.plot);
dev.off()
```


Venn diagriam for downregulated genes at T3
```{r}
#ven diagram

venn.plot <- draw.pairwise.venn(
	area1 = size_T3_sus_down,
	area2 = size_T3_tol_down,
	cross.area = size_T3_down,
	category = c("Drought Susceptible","Drought Tolerant "),
	fill = c("blue", "red"),
	euler.d = TRUE, 
	scaled = TRUE,
	lty = "blank",
	cex = 2,
	cat.dist = -0.02,
	cat.cex = 1.5,
	ext.line.lty = "dashed"
	);


venn.plot <- grid.arrange(gTree(children=venn.plot), top=textGrob("Downregulated Genes at Recovery", gp=gpar(fontsize=25,font=8)) )

# Writing to file
tiff(filename = "Downregulated Genes T3 Root Venn2_diagram.tiff");
grid.draw(venn.plot);
dev.off()
```

Drought Susceptible
genes present 
Comparing upregulated genes in all conditions for Drought Susceptible
```{r}
#Tables and sizes
#T1_sus_up
#size_T1_sus_up 
head(size_T1_sus_up)

#T2_sus_up
#size_T2_sus_up 
head(size_T2_sus_up)

#T3_sus_up
#size_T3_sus_up 
head(size_T3_sus_up)


# T1 vs T2
#list of genes shared between conditions
T1vsT2_up_s = intersect(T1_sus_up$susR_T1_Up_genes, T2_sus_up$susR_T2_Up_genes)
write.csv(data.frame(T1vsT2_up_s), file= "Upregulated genes in both T1 and T2 sRoot.csv", row.names = FALSE)
#number of genes shared between each set
size_T1vsT2_up_s <- length(T1vsT2_up_s)
head(size_T1vsT2_up_s)

# T2 vs T3
#list of genes shared between conditions
T2vsT3_up_s = intersect(T2_sus_up$susR_T2_Up_genes, T3_sus_up$susR_T3_Up_genes)
write.csv(data.frame(T2vsT3_up_s), file= "Upregulated genes in both T2 and T3 sRoot.csv", row.names = FALSE)
#number of genes shared between each set
size_T2vsT3_up_s <- length(T2vsT3_up_s)
head(size_T2vsT3_up_s)


# T1 vs T3
#list of genes shared between conditions
T1vsT3_up_s = intersect(T1_sus_up$susR_T1_Up_genes, T3_sus_up$susR_T3_Up_genes)
write.csv(data.frame(T1vsT3_up_s), file= "Upregulated genes in both T1 and T3 sRoot.csv", row.names = FALSE)
#number of genes shared between each set
size_T1vsT3_up_s <- length(T1vsT3_up_s)
head(size_T1vsT3_up_s)


# T1 vs T2 vs T3
T1vsT2vsT3_up_s = intersect(T1_sus_up$susR_T1_Up_genes, T2_sus_up$susR_T2_Up_genes) 
T1vsT2vsT3_up_s = intersect(T1vsT2vsT3_up_s ,T3_sus_up$susR_T3_Up_genes)
write.csv(data.frame(T1vsT2vsT3_up_s), file= "Downregulated genes in all time.points sRoot.csv", row.names = FALSE)
#number of genes shared between all conditions
size_T1vsT2vsT3_up_s <- length(T1vsT2vsT3_up_s)
head(size_T1vsT2vsT3_up_s)

```

Venn diagram for upregulated genes in all conditions for Drought Susceptible
```{r}
venn.plot <- draw.triple.venn(
	area1 = size_T1_sus_up,
	area2 = size_T2_sus_up,
	area3 = size_T3_sus_up,
	n12 = size_T1vsT2_up_s,
	n23 = size_T2vsT3_up_s,
	n13 = size_T1vsT3_up_s,
	n123 = size_T1vsT2vsT3_up_s,
	category = c("Early 
	Response", 
	  "Late 
	 Response", "Recovery"),
	euler.d = TRUE, 
	scaled = TRUE,
	fill = c("blue", "red", "green"),
	lty = "blank",
	cex = 2,
	cat.dist = 0.1,
	cat.cex = 1.5,
	cat.col = c("blue", "red", "green")
	);

venn.plot <- grid.arrange(gTree(children=venn.plot), top=textGrob("Upregulated Genes in Drought Susceptible Potatoes", gp=gpar(fontsize=22,font=8)) )

# Writing to file
tiff(filename = "Suceptible upregulated genes Root Venn3_diagram.tiff");
grid.draw(venn.plot);
dev.off()

```
Comparing Downregulated genes in all conditions for Drought Susceptible
```{r}
#Tables and sizes
#T1_sus_down
#size_T1_sus_down 
head(size_T1_sus_down)

#T2_sus_down
#size_T2_sus_down 
head(size_T2_sus_down)

#T3_sus_down
#size_T3_sus_down 
head(size_T3_sus_down)


# T1 vs T2
#list of genes shared between conditions
T1vsT2_down_s = intersect(T1_sus_down$susR_T1_Down_genes, T2_sus_down$susR_T2_Down_genes)
write.csv(data.frame(T1vsT2_down_s), file= "downregulated genes in both T1 and T2 sRoot.csv", row.names = FALSE)
#number of genes shared between each set
size_T1vsT2_down_s <- length(T1vsT2_down_s)
head(size_T1vsT2_down_s)

# T2 vs T3
#list of genes shared between conditions
T2vsT3_down_s = intersect(T2_sus_down$susR_T2_Down_genes, T3_sus_down$susR_T3_Down_genes)
write.csv(data.frame(T2vsT3_down_s), file= "downregulated genes in both T2 and T3 sRoot.csv", row.names = FALSE)
#number of genes shared between each set
size_T2vsT3_down_s <- length(T2vsT3_down_s)
head(size_T2vsT3_down_s)


# T1 vs T3
#list of genes shared between conditions
T1vsT3_down_s = intersect(T1_sus_down$susR_T1_Down_genes, T3_sus_down$susR_T3_Down_genes)
write.csv(data.frame(T1vsT3_down_s), file= "downregulated genes in both T1 and T3 sRoot.csv", row.names = FALSE)
#number of genes shared between each set
size_T1vsT3_down_s <- length(T1vsT3_down_s)
head(size_T1vsT3_down_s)


# T1 vs T2 vs T3
T1vsT2vsT3_down_s = intersect(T1_sus_down$susR_T1_Down_genes, T2_sus_down$susR_T2_Down_genes) 
T1vsT2vsT3_down_s = intersect(T1vsT2vsT3_down_s ,T3_sus_down$susR_T3_Down_genes)
write.csv(data.frame(T1vsT2vsT3_down_s), file= "Downregulated genes in all time.points sRoot.csv", row.names = FALSE)
#number of genes shared between all conditions
size_T1vsT2vsT3_down_s <- length(T1vsT2vsT3_down_s)
head(size_T1vsT2vsT3_down_s)
```

Venn diagram for downregulated genes in all conditions for Drought Susceptible
```{r}

venn.plot <- draw.triple.venn(
	area1 = size_T1_sus_down,
	area2 = size_T2_sus_down,
	area3 = size_T3_sus_down,
	n12 = size_T1vsT2_down_s,
	n23 = size_T2vsT3_down_s,
	n13 = size_T1vsT3_down_s,
	n123 = size_T1vsT2vsT3_down_s,
	category = c("Early 
	Response", 
	  "Late 
	 Response", "Recovery"),
	euler.d = TRUE, 
	scaled = TRUE,
	fill = c("blue", "red", "green"),
	lty = "blank",
	cex = 2,
	cat.dist = 0.05,
	cat.cex = 1.5,
	cat.col = c("blue", "red", "green")
	);

venn.plot <- grid.arrange(gTree(children=venn.plot), top=textGrob("Downregulated Genes in Drought Susceptible Potatoes", gp=gpar(fontsize=22,font=8)) )

# Writing to file
tiff(filename = "Suceptible downregulated genes Root Venn3_diagram.tiff");
grid.draw(venn.plot);
dev.off()

```



Drought Tolerant 
genes present 

Comparing upregulated genes in all conditions for Drought Tolerant 
```{r}
#Tables and sizes
#T1_tol_up
#size_T1_tol_up 
head(size_T1_tol_up)

#T2_tol_up
#size_T2_tol_up 
head(size_T2_tol_up)

#T3_tol_up
#size_T3_tol_up 
head(size_T3_tol_up)


# T1 vs T2
#list of genes shared between conditions
T1vsT2_up_t = intersect(T1_tol_up$tolR_T1_Up_genes, T2_tol_up$tolR_T2_Up_genes)
write.csv(data.frame(T1vsT2_up_t), file= "Upregulated genes in both T1 and T2 tRoot.csv", row.names = FALSE)
#number of genes shared between each set
size_T1vsT2_up_t <- length(T1vsT2_up_t)
head(size_T1vsT2_up_t)

# T2 vs T3
#list of genes shared between conditions
T2vsT3_up_t = intersect(T2_tol_up$tolR_T2_Up_genes, T3_tol_up$tolR_T3_Up_genes)
write.csv(data.frame(T2vsT3_up_t), file= "Upregulated genes in both T2 and T3 tRoot.csv", row.names = FALSE)
#number of genes shared between each set
size_T2vsT3_up_t <- length(T2vsT3_up_t)
head(size_T2vsT3_up_t)


# T1 vs T3
#list of genes shared between conditions
T1vsT3_up_t = intersect(T1_tol_up$tolR_T1_Up_genes, T3_tol_up$tolR_T3_Up_genes)
write.csv(data.frame(T1vsT3_up_t), file= "Upregulated genes in both T1 and T3 tRoot.csv", row.names = FALSE)
#number of genes shared between each set
size_T1vsT3_up_t <- length(T1vsT3_up_t)
head(size_T1vsT3_up_t)


# T1 vs T2 vs T3
T1vsT2vsT3_up_t = intersect(T1_tol_up$tolR_T1_Up_genes, T2_tol_up$tolR_T2_Up_genes) 
T1vsT2vsT3_up_t = intersect(T1vsT2vsT3_up_t ,T3_tol_up$tolR_T3_Up_genes)
write.csv(data.frame(T1vsT2vsT3_up_t), file= "Downregulated genes in all time.points tRoot.csv", row.names = FALSE)
#number of genes shared between all conditions
size_T1vsT2vsT3_up_t <- length(T1vsT2vsT3_up_t)
head(size_T1vsT2vsT3_up_t)

```

Venn diagram for upregulated genes in all conditions for Drought Tolerant 
```{r}
venn.plot <- draw.triple.venn(
	area1 = size_T1_tol_up,
	area2 = size_T2_tol_up,
	area3 = size_T3_tol_up,
	n12 = size_T1vsT2_up_t,
	n23 = size_T2vsT3_up_t,
	n13 = size_T1vsT3_up_t,
	n123 = size_T1vsT2vsT3_up_t,
	category = c("Early 
	Response", 
	  "Late 
	 Response", "Recovery"),
	euler.d = TRUE, 
	scaled = TRUE,
	fill = c("blue", "red", "green"),
	lty = "blank",
	cex = 2,
	cat.dist = 0.1,
	cat.cex = 1.5,
	cat.col = c("blue", "red", "green")
	);

venn.plot <- grid.arrange(gTree(children=venn.plot), top=textGrob("Upregulated Genes in Drought Tolerant Potatoes", gp=gpar(fontsize=22,font=8)) )

# Writing to file
tiff(filename = "Tolerant upregulated genes Root Venn3_diagram.tiff");
grid.draw(venn.plot);
dev.off()

```
Comparing Downregulated genes in all conditions for Drought Tolerant 
```{r}
#Tables and sizes
#T1_tol_down
#size_T1_tol_down 
head(size_T1_tol_down)

#T2_tol_down
#size_T2_tol_down 
head(size_T2_tol_down)

#T3_tol_down
#size_T3_tol_down 
head(size_T3_tol_down)


# T1 vs T2
#list of genes shared between conditions
T1vsT2_down_t = intersect(T1_tol_down$tolR_T1_Down_genes, T2_tol_down$tolR_T2_Down_genes)
write.csv(data.frame(T1vsT2_down_t), file= "downregulated genes in both T1 and T2 tRoot.csv", row.names = FALSE)
#number of genes shared between each set
size_T1vsT2_down_t <- length(T1vsT2_down_t)
head(size_T1vsT2_down_t)

# T2 vs T3
#list of genes shared between conditions
T2vsT3_down_t = intersect(T2_tol_down$tolR_T2_Down_genes, T3_tol_down$tolR_T3_Down_genes)
write.csv(data.frame(T2vsT3_down_t), file= "downregulated genes in both T2 and T3 tRoot.csv", row.names = FALSE)
#number of genes shared between each set
size_T2vsT3_down_t <- length(T2vsT3_down_t)
head(size_T2vsT3_down_t)


# T1 vs T3
#list of genes shared between conditions
T1vsT3_down_t = intersect(T1_tol_down$tolR_T1_Down_genes, T3_tol_down$tolR_T3_Down_genes)
write.csv(data.frame(T1vsT3_down_t), file= "downregulated genes in both T1 and T3 tRoot.csv", row.names = FALSE)
#number of genes shared between each set
size_T1vsT3_down_t <- length(T1vsT3_down_t)
head(size_T1vsT3_down_t)


# T1 vs T2 vs T3
T1vsT2vsT3_down_t = intersect(T1_tol_down$tolR_T1_Down_genes, T2_tol_down$tolR_T2_Down_genes) 
T1vsT2vsT3_down_t = intersect(T1vsT2vsT3_down_t ,T3_tol_down$tolR_T3_Down_genes)
write.csv(data.frame(T1vsT2vsT3_down_t), file= "Downregulated genes in all time.points tRoot.csv", row.names = FALSE)
#number of genes shared between all conditions
size_T1vsT2vsT3_down_t <- length(T1vsT2vsT3_down_t)
head(size_T1vsT2vsT3_down_t)
```

Venn diagram for downregulated genes in all conditions for Drought Tolerant 
```{r}

venn.plot <- draw.triple.venn(
	area1 = size_T1_tol_down,
	area2 = size_T2_tol_down,
	area3 = size_T3_tol_down,
	n12 = size_T1vsT2_down_t,
	n23 = size_T2vsT3_down_t,
	n13 = size_T1vsT3_down_t,
	n123 = size_T1vsT2vsT3_down_t,
	category = c("Early 
	Response", 
	  "Late 
	 Response", "Recovery"),
	euler.d = TRUE, 
	scaled = TRUE,
	fill = c("blue", "red", "green"),
	lty = "blank",
	cex = 2,
	cat.dist = 0.1,
	cat.cex = 1.5,
	cat.col = c("blue", "red", "green")
	);

venn.plot <- grid.arrange(gTree(children=venn.plot), top=textGrob("Downregulated Genes in Drought Tolerant Potatoes", gp=gpar(fontsize=22,font=8)) )

# Writing to file
tiff(filename = "Tolerant downregulated genes Root Venn3_diagram.tiff");
grid.draw(venn.plot);
dev.off()

```

Bar chart for number of Upregulated genes for susceptible and tolerant for all conditions
```{r}
#Create Bar chart

Variety <- c("Drought Susceptible", "Drought Susceptible","Drought Susceptible", "Drought Tolarent", "Drought Tolarent", "Drought Tolarent")


#gene count
Gene_Count <- c(size_T1_sus_up, size_T2_sus_up, size_T3_sus_up, size_T1_tol_up, size_T2_tol_up, size_T3_tol_up) 
              

Condition <- c("Early_Response S", "Late_Response S", "Recovery S","Early_Response T", "Late_Response T", "Recovery T")



Up_table <- data.frame(Condition, Gene_Count, Variety)

as.factor(Up_table$Variety)


# Creating plot
Up_Barchart <- ggplot(Up_table, aes(x = Up_table$Condition, y = Up_table$Gene_Count)) +
  geom_col(aes(fill = Up_table$Variety)) +
  theme(axis.text.x = element_text(angle = 50, vjust = 1.0, hjust=1)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  labs(x = "Time Point", y = "Number of Genes", fill = "Condition", title = "Upregulated Gene Expression in Potatoes") +
  theme(plot.title = element_text(hjust = 0.5) 
          
)
#save plot
png("Upregulated Differential Gene Expression Barchart Root.png")
Up_Barchart
dev.off()
Up_Barchart

Condition <-c("Early_Response", "Late_Response", "Recovery","Early_Response", "Late_Response", "Recovery")
Up_table <- data.frame(Condition, Gene_Count, Variety)

as.factor(Up_table$Variety)
g <- ggplot(Up_table, aes(x = Up_table$Condition, y = Up_table$Gene_Count, fill = Up_table$Variety)) +
    labs(x = "Time Point", y = "Number of Genes", fill = "Condition", title = "Upregulated Gene Expression in Potatoes") +
  geom_line() +
  geom_point(size = 4, shape = 21)  # Also use a point with a color fill

#save plot
png("Upregulated Differential Gene Expression line Root.png")
g
dev.off()
g

```
Bar chart for number of Downregulated genes for susceptible and tolerant for all conditions
```{r}
#Create Bar chart

Variety <- c("Drought Susceptible", "Drought Susceptible","Drought Susceptible", "Drought Tolarent", "Drought Tolarent", "Drought Tolarent")


#gene count
Gene_Count <- c(size_T1_sus_down, size_T2_sus_down, size_T3_sus_down, size_T1_tol_down, size_T2_tol_down, size_T3_tol_down) 
              

Condition <- c("Early_Response S", "Late_Response S", "Recovery S","Early_Response T", "Late_Response T", "Recovery T")



Up_table <- data.frame(Condition, Gene_Count, Variety)

as.factor(Up_table$Variety)


# Creating plot
Down_Barchart <- ggplot(Up_table, aes(x = Up_table$Condition, y = Up_table$Gene_Count)) +
  geom_col(aes(fill = Up_table$Variety)) +
  theme(axis.text.x = element_text(angle = 50, vjust = 1.0, hjust=1)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  labs(x = "Time Point", y = "Number of Genes", fill = "Condition", title = "Downregulated Gene Expression in Potatoes") +
  theme(plot.title = element_text(hjust = 0.5) 
          
)

#save plot
png("Downregulated Differential Gene Expression Barchart Root.png")
Down_Barchart
dev.off()
Down_Barchart

Condition <-c("Early_Response", "Late_Response", "Recovery","Early_Response", "Late_Response", "Recovery")
Up_table <- data.frame(Condition, Gene_Count, Variety)

as.factor(Up_table$Variety)
g <- ggplot(Up_table, aes(x = Up_table$Condition, y = Up_table$Gene_Count, fill = Up_table$Variety)) +
    labs(x = "Time Point", y = "Number of Genes", fill = "Condition", title = "Downregulated Gene Expression in Potatoes") +
  geom_line() +
  geom_point(size = 4, shape = 21)  # Also use a point with a color fill

#save plot
png("Downregulated Differential Gene Expression line Root.png")
g
dev.off()
g
```
```


