---
title: "Data overview and visualisation"
author: "Nacima Abdi"
date: "18/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Perform initial quality control 



```{r,warning = FALSE}
#########################################
###   Load the required libraries     ###
#########################################

#Working directory
#setwd("Z:/NA_AreaOnly/Results/Analysis")

# DESeq2 library

library("DESeq2")
# Library for plots
library("ggplot2")
library(RColorBrewer)
library(dplyr)
library("plotly")
# Library for heatmaps
library("pheatmap")


```

Load raw gene-level counts and sample data and prepare for analysis.

```{r}
###   Read the counts from a file   ###
#######################################

# Read the table of counts into a data frame, 
count_table <- as.matrix(read.csv("counts.csv",  row.names="GeneID"))
# Check the top few rows of the data frame
#head(count_table)

# Dimensions of the count table
dim(count_table)
#import sample meta data by coping column D-I from excel 
sample_data <- read.table(file = "clipboard", 
                      sep = "\t", header=TRUE)
#rearrange sample data so column names for count_table matches with library name
colnames(sample_data)
col_order <- c("Library.name", "Variety", "Tissue", "Time.point", "Replicate", "Sample.code")
sample_data <- sample_data[, col_order]
sampleInfo<- as.data.frame(sample_data)
#rename column
names(sampleInfo)[names(sampleInfo) == "Library.name"] <- "Sample"
#order sample data by sample name
sampleInfo <- sampleInfo[order(sampleInfo$Sample),]

rownames(sampleInfo) <- sampleInfo$Sample
sampleID <- sampleInfo$Sample

#save SampleInfo
#wite.csv(sampleInfo,file="sampleInfo.csv",quote=FALSE,row.names=FALSE)


#rename Column name to match sampleID

count_name <- colnames(count_table)
count_names <- c("Length", sampleID)
count_matrix <- count_table
colnames(count_matrix) <- count_names
#delete length column
count_matrix = subset(count_matrix, select = -c(Length))
#save count matrix
#wite.csv(count_matrix,file="count_matrix.csv",row.names=TRUE)

#set species, tissue type and time point as factor
as.factor(sampleInfo$Variety)
as.factor(sampleInfo$Tissue)
as.factor(sampleInfo$Time.point)

```

Removing genes that have low count
```{r}
#################################################################
### Remove genes that have less than 10 reads mapping to them ###
#################################################################
#check dimensions
dim(count_matrix)

#set threshold to 10 
lowcoverage_threshold <- 10

# Index of genes that the maximal count across the samples is 10 or less
lowcoverage<-(rowMax(as.matrix(count_matrix)) <= lowcoverage_threshold)

# Check the numbers
table(lowcoverage)

#Filter out low coverage genes by redefining the count table
count_matrix<-count_matrix[!lowcoverage,]
dim(count_matrix)
```

Using 'count_matrix' and 'sampleInfo', construct DESeqDataSet object

```{r}
dds <- DESeqDataSetFromMatrix(countData = count_matrix, colData = sampleInfo, design = ~ Time.point + Variety + Tissue)

```
VST Transformation of raw reads
blind argument: whether the transformation should be blind to the sample information specified by the design formula. blind=True to perform sample quality assurance. 

```{r}
#vst transform
vsd <- vst(dds, blind = TRUE)
#save vst transformed matrix
#write.csv(assay(vsd), file="vst_transform.csv",quote=FALSE,row.names=TRUE)
```


Assess overall similarity between samples with Heatmap of the sample-to-sample Euclidean distances

```{r,fig.asp = 0.9, fig.width = 25, fig.align="center"}

#transpose data
sampleDists <- dist(t(assay(vsd)), method = "euclidean")
sampleDistMatrix <- as.matrix(sampleDists)

#set annotations
colors <- colorRampPalette(rev(brewer.pal(9, "YlGnBu")))(255)
colour_time <-  brewer.pal(4, "Set1")
names(colour_time) <- unique(sampleInfo$Time.point)
colour_variety<- c('Susceptible: Wila Huaka CIP code: 703248' = "darksalmon", 'Tolerant: Negrita CIP code:703671' = "darkmagenta")
names(colour_variety) <- unique(sampleInfo$Variety)
colour_tissue <- c(Leaf = "green", Root = "red")
names(colour_tissue) <- unique(sampleInfo$Tissue)
colours_anno <- list(Variety = colour_variety, Time = colour_time, Tissue = colour_tissue)



#Generate sample-to-sample Euclidean distance heatmap 


# plot
heatmap_sample <- pheatmap(sampleDistMatrix,
                   clustering_distance_rows=sampleDists,
                   clustering_distance_cols=sampleDists,
                   clustering_method = "average",
                   treeheight_row = 100,
                   treeheight_col = 100,
                   annotation_col = sampleInfo[,2:4],
                   annotation_row = sampleInfo[,2:4],
                   annotation_colors = colours_anno,
                   show_rownames = TRUE,
                   show_colnames = TRUE,
                   main = "Sample to Sample Euclidean Distances Heatmap",
                   cellwidth = 20,
                   cellheight = 20,
                   cutree_rows = 2,
                   cutree_cols = 2,
                   col=colors,
                   fontsize = 20,
                   border_color = "black")

#save plot
#png("Sample to Sample Euclidean Distances Heatmap.png",  width = 1900, height = 1500)
#heatmap_sample
#dev.off()

vsd_cor <- cor(assay(vsd))

# heatmap
phm_cor <- pheatmap(vsd_cor,
                    clustering_method = "average",
                    annotation_col = sampleInfo[,2:4],
                    annotation_row = sampleInfo[,2:4],
                    treeheight_row = 80,
                    treeheight_col = 80,
                    annotation_colors = colours_anno,
                    show_rownames = TRUE,
                    show_colnames = TRUE,
                    cellwidth = 15,
                    cellheight = 15,
                    cutree_rows = 2,
                    cutree_cols = 2,
                    col=rev(colors),
                    fontsize = 12,
                    border_color = "black")
```





Sample CO which is Susceptible leaf sample at late response (SL2-3) is concerning as it is closer to root samples than the rest even though it follows the pattern of late response and recovery.

```{r, fig.asp = 0.8, fig.width = 20, fig.align="center"}
#see if data separates by tissue
plotPCA(vsd, intgroup=c("Tissue"))

#see if data separates by time.point
plotPCA(vsd, intgroup=c("Time.point"))

#see if data separates by Variety
plotPCA(vsd, intgroup=c( "Variety"))

        
pca_data <- plotPCA(vsd, intgroup = c("Tissue", "Time.point", "Variety"), ntop=2000,  returnData = TRUE)
percentVar <- round(100 * attr(pca_data, "percentVar"))
PVA_all <- ggplot(pca_data, aes(PC1, PC2, color=Time.point, shape=Variety)) +
  geom_point(size=5) +
  theme(text = element_text(size=45)) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed()


#save plot
png("Sample PCA .png", width = 1900, height = 1000)
PVA_all
dev.off()

       
```
Biggest variance is due to tissue type explained by pc1  so in order to identify the effect of variety on DGE, I will separate datasets for tissue type and run DGE analysis separately. 

PC2 captures variance due to Time.point



```{r}
#More fancier pca plot
# -----------------------------------------------------
# 4. Define PCA plot
# -----------------------------------------------------
pca.plot <- function(read.counts, classes, 
                     comps = c(1, 2), ntop = min(500, nrow(read.counts)), standard = T,
                     col = c('lightblue', 'orange', 'MediumVioletRed', 'SpringGreen', 'red', 'blue', 'pink', 'black', 'purple', 'maroon', 'brown', 'cadetblue', 'chocolate', 'darkolivegreen', 'firebrick', 'gold', 'darkred')){
  top_index <- order(apply(read.counts, 1, var), decreasing = TRUE)[1:ntop]
  pca <- prcomp(scale(t(read.counts[top_index,]), center = standard, scale = standard))
  prp <- pca$sdev^2 * 100 / sum(pca$sdev^2)
  
  pca_comps <- pca$x[,comps]
  prp_comps <- round(prp[comps], 2)
  
  df <- data.frame(pc1 = pca_comps[,1], pc2 = pca_comps[,2], condition = classes)
  p  <- ggplot(df, aes(x = pc1, y = pc2, color = condition)) + 
    geom_point(size = 3) + 
    labs(title = paste0('Principal Component Analysis - Axes ', comps[1] , ' and ', comps[2]), 
         x = paste0('PC', comps[1], ' (', prp_comps[1], '%)'), 
         y = paste0('PC', comps[2], ' (', prp_comps[2], '%)')) + 
    geom_text(label = colnames(read.counts), vjust = 0, nudge_y = 1) +
    scale_color_manual(values = col)
  return(p)
}
# -----------------------------------------------------
# Plot PCA on the read counts
# -----------------------------------------------------
group_tissue <- vsd$Tissue
group_variety<- vsd$Variety
group_time.point<- vsd$Time.point

pca_tissue <- pca.plot(assay(vsd), group_tissue, comps = c(1,2), ntop = 2000)
ggplotly(pca_tissue)


pca_variety <- pca.plot(assay(vsd), group_variety, comps = c(1,2), ntop = 2000)
ggplotly(pca_variety)


pca_time.point<- pca.plot(assay(vsd), group_time.point, comps = c(1,2), ntop = 2000)
ggplotly(pca_time.point)



```
PC3 and PC4
Performing PCA
```{r}
# Create a matrix from our table of counts
pca_countmatrix <- assay(vsd)
  
  # transpose the matrix so that rows = samples and columns = variables
  pca_countmatrix<-t(pca_countmatrix)

# Perform the PCA
PCA <- prcomp(pca_countmatrix)



#plot PC3 and PC4
pca_data <- plotPCA(vsd, intgroup = c("Tissue", "Time.point", "Variety"),  returnData = TRUE)
percentVar <- round(100 * attr(pca_data, "percentVar"))
ggplot(pca_data, aes(PC3, PC4, color=Time.point, shape=Variety)) +
  geom_point(size=3) +
  xlab(paste0("PC3: ",percentVar[3],"% variance")) +
  ylab(paste0("PC4: ",percentVar[4],"% variance")) + 
  coord_fixed()
```


