---
title: "Leaf analysis"
author: "Nacima Abdi"
date: "29/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r, message = FALSE, warning = FALSE}
#library 
library(goProfiles)
library(tidyverse)
library(ggvenn)
library(VennDiagram)

```





Examine KEGG and GO enrichment under normal conditions

```{r, message = FALSE, warning = FALSE}
#DEG between susceptible and tolerant

#load data

Control_up_genes = read.csv("Upregulated_genes_control_Leaf.csv")

Control_down_genes = read.csv("Downregulated_genes_Control_Leaf.csv")



```

genes common in both variety at each condition
```{r}
#t1
#up
T1_sus_up <- read.csv("Upregulated_genes_susceptible_T1_Leaf.csv")

T1_tol_up <-read.csv("Upregulated_genes_tolerant_T1_Leaf.csv")

#match 
T1_up_genes_overlap = intersect(T1_sus_up$susL_T1_Up_genes, T1_tol_up$tolL_T1_Up_genes)
write.csv(data.frame(T1_up_genes_overlap), file= "Upregulated_genes_common_T1_Leaf.csv", row.names = FALSE)

#down
T1_sus_down <- read.csv("Downregulated_genes_susceptible_T1_Leaf.csv")

T1_tol_down <-read.csv("Downregulated_genes_tolerant_T1_Leaf.csv")

#match 
T1_down_genes_overlap = intersect(T1_sus_down$susL_T1_Down_genes, T1_tol_down$tolL_T1_Down_genes)
write.csv(data.frame(T1_down_genes_overlap), file= "Downregulated_genes_common_T1_Leaf.csv", row.names = FALSE)
```

```{r}
#2
#up
T2_sus_up <- read.csv("Upregulated_genes_susceptible_T2_Leaf.csv")

T2_tol_up <-read.csv("Upregulated_genes_tolerant_T2_Leaf.csv")

#match 
T2_up_genes_overlap = intersect(T2_sus_up$susL_T2_Up_genes, T2_tol_up$tolL_T2_Up_genes)
write.csv(data.frame(T2_up_genes_overlap), file= "Upregulated_genes_common_T2_Leaf.csv", row.names = FALSE)

#down
T2_sus_down <- read.csv("Downregulated_genes_susceptible_T2_Leaf.csv")

T2_tol_down <-read.csv("Downregulated_genes_tolerant_T2_Leaf.csv")

#match 
T2_down_genes_overlap = intersect(T2_sus_down$susL_T2_Down_genes, T2_tol_down$tolL_T2_Down_genes)
write.csv(data.frame(T2_down_genes_overlap), file= "Downregulated_genes_common_T2_Leaf.csv", row.names = FALSE)

```

```{r}
#t3
#up
T3_sus_up <- read.csv("Upregulated_genes_susceptible_T3_Leaf.csv")

T3_tol_up <-read.csv("Upregulated_genes_tolerant_T3_Leaf.csv")

#match 
T3_up_genes_overlap = intersect(T3_sus_up$susL_T3_Up_genes, T3_tol_up$tolL_T3_Up_genes)
write.csv(data.frame(T3_up_genes_overlap), file= "Upregulated_genes_common_T3_Leaf.csv", row.names = FALSE)

#down
T3_sus_down <- read.csv("Downregulated_genes_susceptible_T3_Leaf.csv")

T3_tol_down <-read.csv("Downregulated_genes_tolerant_T3_Leaf.csv")

#match 
T3_down_genes_overlap = intersect(T3_sus_down$susL_T3_Down_genes, T3_tol_down$tolL_T3_Down_genes)
write.csv(data.frame(T3_down_genes_overlap), file= "Downregulated_genes_common_T3_Leaf.csv", row.names = FALSE)




```


genes unique to varity at each condition
```{r}

#susceptible 
#load data
#T1
T1_seq_s <- read.csv("DEG_sig_results_susceptible_T1_Leaf.csv")
T1_seq_s_genes <- T1_seq_s$X

#T2
T2_seq_s <- read.csv("DEG_sig_results_susceptible_T2_Leaf.csv")
#T1
T3_seq_s <- read.csv("DEG_sig_results_susceptible_T3_Leaf.csv")

#tolerant
#load data
#T1
T1_seq_t <- read.csv("DEG_sig_results_tolerant_T1_Leaf.csv")
#T2
T2_seq_t <- read.csv("DEG_sig_results_tolerant_T2_Leaf.csv")
#T1
T3_seq_t <- read.csv("DEG_sig_results_tolerant_T3_Leaf.csv")

```



Susceptible
genes present 

Venn diagram for upregulated genes in all conditions for sus

```{r}
#data
#number of genes in each set
size_T1_sus_up <- nrow(T1_sus_up)
size_T2_sus_up <- nrow(T2_sus_up)
size_T3_sus_up <- nrow(T3_sus_up)
#list of genes shared between conditions
T1vsT2_up_genes_overlap_s = intersect(T1_sus_up$susL_T1_Up_genes, T2_sus_up$susL_T2_Up_genes)
write.csv(data.frame(T1vsT2_up_genes_overlap_s), file= "Upregulated_genes in all time.points__tRoot.csv", row.names = FALSE)



T1vsT3_up_genes_overlap_s = intersect(T1_sus_up$susL_T1_Up_genes, T3_sus_up$susL_T3_Up_genes)
T2vsT3_up_genes_overlap_s = intersect(T2_sus_up$susL_T2_Up_genes, T3_sus_up$susL_T3_Up_genes)
T1vsT2vsT3_up_genes_overlap_s = intersect(T1vsT2_up_genes_overlap_s, T3_sus_up$susL_T3_Up_genes)

#number of genes shared between each set
size_T1vsT2_up_s <- length(T1vsT2_up_genes_overlap_s)
size_T1vsT3_up_s <- length(T1vsT3_up_genes_overlap_s)
size_T2vsT3_up_s <- length(T2vsT3_up_genes_overlap_s)
size_T1vsT2vsT3_up_s <- length(T1vsT2vsT3_up_genes_overlap_s)

venn.plot <- draw.triple.venn(
	area1 = size_T1_sus_up,
	area2 = size_T2_sus_up,
	area3 = size_T3_sus_up,
	n12 = size_T1vsT2_up_s,
	n23 = size_T2vsT3_up_s,
	n13 = size_T1vsT3_up_s,
	n123 = size_T1vsT2vsT3_up_s,
	category = c("Early Response", "Late Response", "Recovery"),
	euler.d = TRUE, 
	scaled = TRUE,
	fill = c("blue", "red", "green"),
	lty = "blank",
	cex = 2,
	cat.dist = 0.08,
	cat.cex = 2,
	cat.col = c("blue", "red", "green")
	);
grid.draw(venn.plot);
grid.newpage();
# Writing to file
tiff(filename = "Suceptible_upregulated genes_Venn3_diagram_Leaf.tiff", width= 800, compression = "lzw");
grid.draw(venn.plot);
dev.off()

```
Venn diagram for downregulated genes in all conditions for sus

```{r}
data
#number of genes in each set
size_T1_sus_down <- nrow(T1_sus_down)
size_T2_sus_down <- nrow(T2_sus_down)
size_T3_sus_down <- nrow(T3_sus_down)
#list of genes shared between conditions
T1vsT2_down_genes_overlap_s = intersect(T1_sus_down$susL_T1_down_genes, T2_sus_down$susL_T2_down_genes)
T1vsT3_down_genes_overlap_s = intersect(T1_sus_down$susL_T1_down_genes, T3_sus_down$susL_T3_down_genes)
T2vsT3_down_genes_overlap_s = intersect(T2_sus_down$susL_T2_down_genes, T3_sus_down$susL_T3_down_genes)
T1vsT2vsT3_down_genes_overlap_s = intersect(T1vsT2_down_genes_overlap_s, T3_sus_down$susL_T3_down_genes)

#number of genes shared between each set
size_T1vsT2_down_s <- length(T1vsT2_down_genes_overlap_s)
size_T1vsT3_down_s <- length(T1vsT3_down_genes_overlap_s)
size_T2vsT3_down_s <- length(T2vsT3_down_genes_overlap_s)
size_T1vsT2vsT3_down_s <- length(T1vsT2vsT3_down_genes_overlap_s)

venn.plot <- draw.triple.venn(
	area1 = size_T1_sus_down,
	area2 = size_T2_sus_down,
	area3 = size_T3_sus_down,
	n12 = size_T1vsT2_down_s,
	n23 = size_T2vsT3_down_s,
	n13 = size_T1vsT3_down_s,
	n123 = size_T1vsT2vsT3_down_s,
	category = c("Early Response", "Late Response", "Recovery"),
	euler.d = TRUE, 
	scaled = TRUE,
	fill = c("blue", "red", "green"),
	lty = "blank",
	cex = 2,
	cat.dist = 0.08,
	cat.cex = 2,
	cat.col = c("blue", "red", "green")
	);
grid.draw(venn.plot);
grid.newpage();
# Writing to file
tiff(filename = "Suceptible_downregulated genes_Venn3_diagram_Leaf.tiff", width= 800, compression = "lzw");
grid.draw(venn.plot);
dev.off()

```
no overlap


Venn diagram for upregulated genes in all conditions for tol

```{r}
#data
#number of genes in each set
size_T1_tol_up <- nrow(T1_tol_up)
head(size_T1_tol_up)
size_T2_tol_up <- nrow(T2_tol_up)
head(size_T2_tol_up)
size_T3_tol_up <- nrow(T3_tol_up)
head(size_T3_tol_up)
#list of genes shared between conditions
T1vsT2_up_genes_overlap_t = intersect(T1_tol_up$tolL_T1_Up_genes, T2_tol_up$tolL_T2_Up_genes)
T1vsT3_up_genes_overlap_t = intersect(T1_tol_up$tolL_T1_Up_genes, T3_tol_up$tolL_T3_Up_genes)
T2vsT3_up_genes_overlap_t = intersect(T2_tol_up$tolL_T2_Up_genes, T3_tol_up$tolL_T3_Up_genes)
T1vsT2vsT3_up_genes_overlap_t = intersect(T1vsT2_up_genes_overlap_s, T1vsT3_up_genes_overlap_t)

#number of genes shared between each set
size_T1vsT2_up_t <- length(T1vsT2_up_genes_overlap_t)
size_T1vsT3_up_t <- length(T1vsT3_up_genes_overlap_t)
size_T2vsT3_up_t <- length(T2vsT3_up_genes_overlap_t)
size_T1vsT2vsT3_up_t <- length(T1vsT2vsT3_up_genes_overlap_t)

#venn.plot <- draw.triple.venn(
#	area1 = size_T1_tol_up,
#	area2 = size_T2_tol_up,
#	area3 = size_T3_tol_up,
#	n12 = size_T1vsT2_up_t, 
#	n23 = size_T2vsT3_up_t,
#	n13 = size_T1vsT3_up_t,
#	n123 = size_T1vsT2vsT3_up_t,
#	category = c("Early Response", "Late Response", "Recovery"),
#	euler.d = TRUE, 
#	scaled = TRUE,
#	fill = c("blue", "red", "green"),
#	lty = "blank",
#	cex = 2,
#	cat.dist = 0.08,
#	cat.cex = 2,
#	cat.col = c("blue", "red", "green")
#	);
#grid.draw(venn.plot);
#grid.newpage();
# Writing to file
#tiff(filename = "Tolarent_upregulated genes_Venn3_diagram_Leaf.tiff", width= 800, compression = "lzw");
#grid.draw(venn.plot);
dev.off()

```

```{r}
data
#number of genes in each set
size_T1_tol_down <- nrow(T1_tol_down)
size_T2_tol_down <- nrow(T2_tol_down)
size_T3_tol_down <- nrow(T3_tol_down)
#list of genes shared between conditions
T1vsT2_down_genes_overlap_t = intersect(T1_tol_down$tolL_T1_down_genes, T2_tol_down$tolL_T2_down_genes)
T1vsT3_down_genes_overlap_t = intersect(T1_tol_down$tolL_T1_down_genes, T3_tol_down$tolL_T3_down_genes)
T2vsT3_down_genes_overlap_t = intersect(T2_tol_down$tolL_T2_down_genes, T3_tol_down$tolL_T3_down_genes)
T1vsT2vsT3_down_genes_overlap_t = intersect(T1vsT2_down_genes_overlap_t, T3_tol_down$tolL_T3_down_genes)

#number of genes shared between each set
size_T1vsT2_down_t <- length(T1vsT2_down_genes_overlap_t)
size_T1vsT3_down_t <- length(T1vsT3_down_genes_overlap_t)
size_T2vsT3_down_t <- length(T2vsT3_down_genes_overlap_t)
size_T1vsT2vsT3_down_t <- length(T1vsT2vsT3_down_genes_overlap_t)

venn.plot <- draw.triple.venn(
	area1 = size_T1_tol_down,
	area2 = size_T2_tol_down,
	area3 = size_T3_tol_down,
	n12 = size_T1vsT2_down_t,
	n23 = size_T2vsT3_down_t,
	n13 = size_T1vsT3_down_t,
	n123 = size_T1vsT2vsT3_down_t,
	category = c("Early Response", "Late Response", "Recovery"),
	euler.d = TRUE, 
	scaled = TRUE,
	fill = c("blue", "red", "green"),
	lty = "blank",
	cex = 2,
	cat.dist = 0.08,
	cat.cex = 2,
	cat.col = c("blue", "red", "green")
	);
grid.draw(venn.plot);
grid.newpage();
# Writing to file
tiff(filename = "Tolarent_downregulated genes_Venn3_diagram_Leaf.tiff", width= 800, compression = "lzw");
grid.draw(venn.plot);
dev.off()


```
Bar chart for number of Upregulated genes for susceptible and tolerant for all conditions
```{r}
#Create Bar chart

Variety <- c("Drought Tolarent", "Drought Tolarent", "Drought Tolarent","Drought Susceptible", "Drought Susceptible","Drought Susceptible")
DGE_Direction <- c("Upregulated", "Upregulated","Upregulated","Upregulated", "Upregulated","Upregulated")

#gene count
Gene_Count <- c(size_T1_sus_up, size_T2_sus_up, size_T3_sus_up, size_T1_tol_up, size_T2_tol_up, size_T3_tol_up) 
                

Condition <- c("Early_Response S", "Late_Response S", "Recovery S","Early_Response T", "Late_Response T", "Recovery T")



Up_table <- data.frame(Condition, DGE_Direction, Gene_Count, Variety)

as.factor(Up_table$Variety)


# Creating plot
#Su_Barchart <- 
ggplot(Up_table, aes(x = Up_table$Condition, y = Up_table$Gene_Count)) +
  geom_col(aes(fill = Up_table$Variety)) +
  theme(axis.text.x = element_text(angle = 50, vjust = 1.0, hjust=1)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  labs(x = "Time Point", y = "Number of Genes", fill = "Condition", title = "Upregulated Differential Gene Expression") +
  theme(plot.title = element_text(hjust = 0.5) 
          
)

Condition <-c("Early_Response", "Late_Response", "Recovery","Early_Response", "Late_Response", "Recovery")
Up_table <- data.frame(Condition, DGE_Direction, Gene_Count, Variety)

as.factor(Up_table$Variety)
ggplot(Up_table, aes(x = Up_table$Condition, y = Up_table$Gene_Count, fill = Up_table$Variety)) +
    labs(x = "Time Point", y = "Number of Genes", fill = "Condition", title = "Upregulated Differential Gene Expression") +
  geom_line() +
  geom_point(size = 4, shape = 21)  # Also use a point with a color fill



```

Bar chart for number of down regulated genes for susceptible and tolerant for all conditions
```{r}

Variety <- c("Drought Tolarent", "Drought Tolarent", "Drought Tolarent","Drought Susceptible", "Drought Susceptible","Drought Susceptible")
DGE_Direction <- c("Downregulated","Downregulated","Downregulated","Downregulated","Downregulated","Downregulated")

#gene count
Gene_Count <- c(size_T1_sus_down, size_T2_sus_down, size_T3_sus_down, size_T1_tol_down, size_T2_tol_down, size_T3_tol_down) 
               

Condition <- c("Early_Response S", "Late_Response S", "Recovery S","Early_Response T", "Late_Response T", "Recovery T")



Up_table <- data.frame(Condition, DGE_Direction, Gene_Count, Variety)

as.factor(Up_table$Variety)


# Creating plot
#Su_Barchart <- 
ggplot(Up_table, aes(x = Up_table$Condition, y = Up_table$Gene_Count)) +
  geom_col(aes(fill = Up_table$Variety)) +
  theme(axis.text.x = element_text(angle = 50, vjust = 1.0, hjust=1)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  labs(x = "Time Point", y = "Number of Genes", fill = "Condition", title = "Upregulated Differential Gene Expression") +
  theme(plot.title = element_text(hjust = 0.5) 
          
)

Condition <-c("Early_Response", "Late_Response", "Recovery","Early_Response", "Late_Response", "Recovery")
Up_table <- data.frame(Condition, DGE_Direction, Gene_Count, Variety)

as.factor(Up_table$Variety)
ggplot(Up_table, aes(x = Up_table$Condition, y = Up_table$Gene_Count, fill = Up_table$Variety)) +
    labs(x = "Time Point", y = "Number of Genes", fill = "Condition", title = "Downregulated Differential Gene Expression") +
  geom_line() +
  geom_point(size = 4, shape = 21)  # Also use a point with a color fill



```